\section{Session Two}\label{sec:session_two}

\subsection{Toolsets}\label{sub:toolsets}

There are many types of toolset:

\subsubsection{Work Item Tracking}\label{ssub:work_item_tracking}

\emph{Work Item Tracking} does what it says.
It tracks different pieces of work and allows you to assign different pieces of work to different people while keeping track of their progress.

\subsubsection{Version Control Systems}\label{ssub:version_control_systems}

Tracks the author and creation date of every change to allow for traceability.
Individuals changes are named \emph{commits} and go onto \emph{branches}.

\subsubsection{Integrated Development Environments}\label{ssub:integrated_development_environments}

IDEs often have many, many features but several exist in almost all of them.

\begin{itemize}
	\item Edit code
	\item Navigate code bases
	\item Integrate with version control and other tools
	\item Debugging (breakpoints, etc.)
	\item Auto formatting and highlighting
	\item ``Impact Analysis'': What does a method call and what calls it?
\end{itemize}

\subsubsection{Software Metrics}\label{ssub:software_metrics}

Software metrics are used to measure different parts of the code base:

\begin{itemize}
	\item Lines of code added, removed, broken, edited, etc.
	\item Cohesion: ``How specific is something''
	\item Coupling: ``How many different things does it call?''
	\item Comment density
	\item Test coverage
\end{itemize}

\subsubsection{Continuous Integration/Continuous Delivery}\label{ssub:continuous_integration_continuous_delivery}

\begin{itemize}
	\item Automated building
	\item Run tests in a production-like environment
	\item Fix broken builds quickly with early feedback to revert changes
	\item Automated deployments (every 2 days, 2 hours, etc.)
	\item Several different environments:
	      \begin{itemize}
		      \item \emph{Local:} On my local machine
		      \item \emph{Development:} Probably broken because programmers may be testing
		      \item \emph{System Integration:} Used to test how the code works with other systems
		      \item \emph{User Acceptance:} Does the code work for users?
		      \item \emph{Performance Testing:} What happens if there's a slow network connection? What about no RAM?
		      \item \emph{Live/Production:} Where the code gets used by end users
	      \end{itemize}
\end{itemize}

\subsubsection{Test Tools}\label{ssub:test_tools}

\begin{itemize}
	\item Track tests to their original requirements
	\item Unit testing individual sections
	\item Performance testing
	\item Track status of tests
\end{itemize}

\subsubsection{Communication}\label{ssub:communication}

\begin{itemize}
	\item Necessary for remote working where developers don't see each other all day.
	\item Might be hard to time meetings if there's more than one country involved.
	\item Could be ``scrum'' or ``stand-up'' meetings.
	\item For meetings, get a list of actions for each person and record any decisions made.
	\item Do meetings affect developer productivity (Americans love \emph{really} big meetings with lots of people that don't need to be there)
\end{itemize}

\subsubsection{Application Documentation}\label{ssub:application_documentation}

\begin{itemize}
	\item Should record project scope (what will/won't be included)
	\item Instructions on how to perform a deploy
	\item User help issues
	\item Issue logs for future reference
	\item Hardware architecture
	\item How to / FAQ
\end{itemize}

\begin{note}
	Many of these tools integrate together to create a seamless experience.
\end{note}
