\section{Session Six}\label{sec:fpse_session_six}

\subsection{Coupling and Cohesion}\label{sub:coupling_and_cohesion}

\paragraph{Coupling}\label{par:coupling}

How interdependent applications, modules or services are on each other.

\paragraph{Cohesion}\label{par:cohesion}

How well units belong together and perform similar business functions.

\subsubsection{Why avoid coupling and cohesion?}\label{ssub:why_avoid_coupling_and_cohesion_}

\begin{itemize}

	\item A database structure change means I have to update my application.
	\item Server configurations could change.
	\item Requires high degree of coordination between teams which can cause delays.
	\item Additional testing needed because of interdependencies.
	\item Additional impact risk.
\end{itemize}

\subsubsection{Why want coupling and cohesion?}\label{ssub:why_want_coupling_and_cohesion_}

\begin{itemize}
	\item The same procedures can be used to get the same data.
	\item Avoids having multiple REST services to let different applications get different data.
	\item To get data from one service, it needs to get data from another first
	      \begin{itemize}
		      \item More network traffic
		      \item More coordination again
		      \item Teams may be too busy to make the changes you need
	      \end{itemize}
\end{itemize}

\subsubsection{Types of Coupling}\label{ssub:types_of_coupling}

\begin{itemize}
	\item \textbf{Data Coupling:} Modules depend on data passed between them.
	\item \textbf{Common Coupling:} Two modules have access to the same data so bad side effects can happen if one changes it without the other knowing.
	\item \textbf{External Coupling:} Modules depend on external software (protocols, external files, hardware devices, etc.).
	\item \textbf{Stamp Coupling:} Modules use the same data structures.
	\item \textbf{Control Coupling:} One modules controls the flow of another by passing data or instructions to it.
	\item \textbf{Subclass Coupling:} A child class is connected to its parent, but the child is not attached to the child.
	\item \textbf{Temporal Coupling:} When two actions are bundled together simply because they happen at the same time.
\end{itemize}

\begin{note}
	\begin{itemize}
		\item[Note:] Seems like absolutely nobody gets this.
	\end{itemize}
\end{note}

\subsubsection{Examples of Cohesion}\label{ssub:examples_of_cohesion}

\paragraph{Monolith}\label{par:monolith}

One application or system doing lots of things with low coupling.
The application could have support for exam papers and research papers.
To add a new committee paper to the system could cause potentially breaking changes.

\paragraph{Separate Systems}\label{par:separate_systems}

Separate systems with low coupling.
You would need different applications for exam papers, research papers and an overall document management system.
Adding a new committee paper here would be simple and just involve the document management system.

\begin{note}
	\begin{itemize}
		\item[Note:] Separate applications with high coupling can become a monolithic application.
	\end{itemize}
\end{note}

\subsubsection{Benefits of Monolithic and Separate Applications}\label{ssub:benefits_of_monolithic_and_separate_applications}

\paragraph{Monolithic (low cohesion)}\label{par:monolithic_low_cohesion_}

\begin{itemize}
	\item Easier to develop on (one big program)
	\item Not networking or other communication needed
	\item Far fewer different technologies can be used
	\item Support is easier to monitor (data is in the same format already and a monolith application can show logging making it easier for users)
	\item Easier to create aggregate reports on all students, trades, sales, etc.
\end{itemize}

\paragraph{Separate Applications (high cohesion)}\label{par:separate_applications_high_cohesion_}

\begin{itemize}
	\item Independent development
	      \begin{itemize}
		      \item This is good since you will have specialists on each program and there will be no inter-team communication.
		      \item This is bad since certain features will need several teams causing delays.
	      \end{itemize}
	\item All applications can be scaled individually (scale some more than others)
	\item Applications can be reused easily (a PDF generator will be used lots of places)
\end{itemize}

\subsubsection{Solutions to Coupling and Cohesion}\label{ssub:solutions_to_coupling_and_cohesion}

\begin{itemize}
	\item \textbf{REST Web Services:} Web requests used to communicate between apps.
	\item \textbf{Messaging:} A queue for tasks, works to limit volume (REST services have to do work \emph{now}).
	\item \textbf{Microservices:} Independent programs to be deployed/scaled separately (communication is still needed here).
	\item \textbf{Feature Switches:} Do work ahead of time for features to be turned on when other teams need it.
\end{itemize}

\paragraph{REST Services}\label{par:rest_services}

\begin{itemize}
	\item Refined in and out parameters
	\item Automated testing
	\item Version numbers for different features
	\item Easy to see who's calling what functions when
	\item Restrict access to certain users
\end{itemize}

\paragraph{Messaging}\label{par:messaging}

\begin{itemize}
	\item Needs a messaging bus set-up
	\item Will need a different separate configuration
	\item Needs to handle sending and receiving data
	\item Needs to parse data (JSON, XML, YAML)
\end{itemize}

\paragraph{Microservices}\label{par:microservices}

\begin{itemize}
	\item High cohesion and low coupling
	\item ``Micro'' doesn't have to mean small
	\item Simpler to understand
	\item Harder to develop on since they are lower level and need to implement communication
\end{itemize}

\paragraph{Feature Switches}\label{par:feature_switches}

\begin{itemize}
	\item Just change a configuration option to enable
	\item Just switch on once other systems need the feature
	\item Lets you enable several features across systems all at once
	\item Still has risks of regular deployment, but is safer
	\item Can introduce technical debt because of extra unneeded configurations (unless removed later)
\end{itemize}
