\section{Session 7}\label{sec:session_7}

\subsection{Coupling and Cohesion Recap}\label{sub:coupling_and_cohesion_recap}

Coupling is a measure of how dependant modules are on each other; does the code require something outside of the application (final executable), module (a file or class) or service (a process available to make API calls, etc.).

\subsubsection{Coupling Types}\label{ssub:coupling_types}

\paragraph{Content Coupling}\label{par:content_coupling}

A modules, file or class uses code from another file module or class. Must be at least on method or variable.

\paragraph{Common Coupling}\label{par:common_coupling}

Several files , classes or modules have access to the same global data (eg.\ more than one class accessing the same database).

\paragraph{External Coupling}\label{par:external_coupling}

Several modules have access to an externally exposed data format, communication protocol or device interface. eg.\ A certain JSON structure.

\paragraph{Control Coupling}\label{par:control_coupling}

A module controls the flow of another module. eg.\ Another module calculates  if someone is bankrupt.

\paragraph{Stamp Coupling}\label{par:stamp_coupling}

Several modules use the same composite data structure.

\paragraph{Data Coupling}\label{par:data_coupling}

Several modules share data through parameters.

\noindent
In order from tightest to loosest coupling:

\begin{enumerate}
	\item Content Coupling
	\item Common Coupling
	\item External Coupling
	\item Control Coupling
	\item Stamp Coupling
	\item Data Coupling
\end{enumerate}

\subsubsection{Cohesion}\label{ssub:cohesion}

Cohesion is how specialised a modules is.

\paragraph{Business Functions}\label{par:business_functions}

A business function is something the end user would expect. eg.\ Creating a deck of cards or dealing a deck of cards.

\paragraph{Technical Functions}\label{par:technical_fucntions}

A technical function is something the program has to do to achieve the business functions. eg.\ A database query or message sending.

\paragraph{High Cohesion}\label{par:high_cohesion}

A very specialised program which likely has high coupling too.

\paragraph{Low Cohesion}\label{par:low_cohesion}

A program that does many, many things which likely has low coupling.
