\section{Session Five}\label{sec:session_five}

\subsection{Subprograms}\label{sub:subprograms}

A \emph{subprogram} is a sequence of statements that can be called more than once (they are named blocks of code).
Subprograms often return a value which allows the subprogram to give feedback.
Sometimes these are called \emph{subroutines}, \emph{procedures}, \emph{functions} or \emph{methods}.
Subprograms temporarily reroute the flow of execution with isolated local variables and control the flow of data with arguments and returns.

\subsubsection{Return}\label{ssub:return}

Gives control back to the calling code from any point inside a function (even half way through).
If return is skipped, execution will end at the end of the code block.

\subsubsection{Call Stack}\label{ssub:call_stack}

Most languages have a stack of calls so that it always knows where code should return to.
When a function is called, it puts the point to return to onto the stack.
When a function returns again, it pops off the function and returns back to the point.

\subsubsection{Parametrisation}\label{ssub:parameterisation}

Functions are only really useful if they can be called with different data.
This is why we parametrise the functions to send data from the calling code.
\emph{Parameters} are the received data.
\emph{Arguments} are passed in.
Parameters are always required, but may be given a default value so they aren't need to be passed in every time.

\subsubsection{Why do we use functions?}\label{ssub:why_do_we_use_functions_}

\paragraph{Decomposition}\label{par:decomposition}

Makes code more readable and manageable.

\paragraph{Isolation}\label{par:isolation}

Isolate variables used in one chunk from those in another.

\paragraph{Abstraction}\label{par:abstraction}

Create a general function that can be used in many places for different problems.

\paragraph{Laziness}\label{par:laziness}

They can eliminate boring, repetitive tasks.
This is sometimes called \emph{generalisation} and is often done through abstraction. % link here

\paragraph{Remove repetition}\label{par:remove_repetition}

Repetition makes it very easy to make copy-paste mistakes and makes code harder to read.

\paragraph{Code sharing}\label{par:code_sharing}

Code in functions is far easier to share with other people if it is in chunks already.

\paragraph{Control the Effects}\label{par:control_the_effects}

Code in functions only affects local variables so there is a smaller risk of it breaking something outside.
